A spiral galaxy's mass distribution uncovered through lensing and dynamics

### 2016-06-13 LensingPaper_draft_v4.01.tex
Corrected some referencing.

%--------------------------------------------
As mentioned above, JAM modelling allows to incorporate an invisible matter component in addition to the luminous matter in the form of an MGE. In Section \ref{sec:results_JAM_NFW} we will include a spherical Navarro-Frenk-White (NFW) DM halo \citep{1996ApJ...462..563N} in the dynamical model. The classical NFW profile has the form
\begin{equation}
\rho_\text{NFW}(r) \propto r^{-1} \left( r+r_s \right)^{-2} \label{eq:NFWprofile}
\end{equation}
and two free parameters, the scale length $r_s$ and a parameter describing the total mass of the halo. We will use $v_\text{200}\NEW{\equiv \sqrt{GM_{200}/r_{200}}}$, which is the circular velocity at the radius $r_\text{200}$ within which the mean density of the halo is 200 times the cosmological critical density $\rho_\text{crit} \equiv (2H^2)/(8\pi G)$, \OLD{i.e.,}\NEW{ and where $M_{200} \equiv M(<r_{200}) \stackrel{!}{=} \frac 43 \pi r_{200}^3 \cdot \rho_\text{crit}(z=0)$.}
\OLD{\begin{eqnarray*}
M_\text{200} &=& M(<r_{200})\\
\frac{M_{200}}{ \frac 43 \pi r_{200}^3} &=& 200 \rho_\text{crit}(z=0) \\
v_\text{200} &=& \sqrt{\frac{GM_{200}}{r_\text{200}}}
\end{eqnarray*}}
with $\rho_\text{crit}(z=0)=1.43 \cdot 10^{-7} M_\odot / \text{pc}^3$ in the WMAP5 cosmology by \citet{WMAP5cosm}. \OLD{How much the mass is concentrated in the center of the NFW halo is given by the concentration parameter
\begin{equation}
c_{200}\equiv r_{200} / r_s. \label{eq:NFW_c}
\end{equation}}
\NEW{The mass concentration of the NFW halo is quantified by $c_{200}\equiv r_{200} / r_s$.}
There is a close relation between the concentration and halo mass in simulations \citep{1996ApJ...462..563N}. \citet{Maccio08} found this relation for the WMAP5 cosmology to be
\begin{equation}
\langle \log c_{200} \rangle (M_{200}) = 0.830 - 0.098 \log \left(h \frac{M_{200}}{10^{12} M_\odot} \right) \label{eq:Maccio08}
\end{equation}
(their equation 10), with a Gaussian scatter of $\sigma_{\log c_{200}} = 0.105$ (their Table A2). \NEW{In Section \Wilma{[TO DO ???]} we will use this relation as prior information to guide the modelling towards realistic NFW halo shapes.}
%--------------------------------------------

@@

%--------------------------------------------
\subsubsection{Best fit JAM modelt} \label{sec:results_JAM_NFW_results}
%--------------------------------------------

@@

\NEW{A fit to all nine available $v_\text{rms}$ data points would force the DM halo to be less massive. But with decreasing DM contribution in the inner regions the model approaches the mass-follows-light model in the center, that we already ruled out in Section \ref{sec:results_JAM_SB_MfL}. Allowing for free $\Upsilon_{*}(R')$, $\beta_z(R')$ and/or non-NFW $\rho_\text{halo}(r)$ profiles would over-fit the data again. Without further priors on the stellar $\Upsilon_{*}(R')$ and/or more data points, we can make only the following statement: A model with constant $\Upsilon_*$, moderately tangential \Wilma{[TO DO: Check]} $\beta_z$ and a spherical and massive DM component with high concentration can explain the central kinematics of J1331 very well, we have however ruled out that in this case the DM halo will follow a NFW halo at both small and large radii.}
%--------------------------------------------

@@

%--------------------------------------------
\subsection{On J1331's possible merger history\OLD{ and modelling failures}}
%--------------------------------------------
\NOTE{[This subsection was moved from a later position to the beginning of this section.]}

@@

\NEW{\emph{(i) Indications for a Salpeter-like IMF in J1331's bulge:}}

@@

\NEW{This rise could be either due to a cuspy DM halo (see Section \Wilma{4.4???}) or a very strong gradient in $\Upsilon_\text{I,*}$ (see Section \Wilma{5.1???}). However, we rule out that a DM cusp is the sole reason because a cuspy and therefore overall massive DM halo does not match the kinematics in J1331's outer regions. We also rule out that a very strong $\Upsilon_\text{I,*}$ gradient is the only reason, because---as mentioned in Section \Wilma{???}---photometry did not reveal the clear existence of a blue population in the very center of J1331's bulge. Also, our modelling attempts allowing for velocity anisotropy (Sections \Wilma{4.3.2} and \Wilma{4.4}) suggest that we do need some moderate $\beta_z<0$ to explain the central $v_\text{rms}$ dip. And lastly---as laid out in the previous paragraph---J1331's central velocity dispersion suggests a more bottom-heavy IMF. Overall it is therefore not very likely that the central regions of J1331 have such a low $\Upsilon_\text{I,*}^\text{chab} \sim 2.5$.}

@@

\NEW{The central $v_\text{rms}$ is most likely a combination of the three different aspects that we investigated separately: We expect a bit of tangential \Wilma{???} velocity anisotropy (Sections \ref{sec:results_JAM_SB_MfL} and \ref{sec:results_JAM_NFW}), a moderate contribution of a roundish DM halo (Sections 4.2.3 \Wilma{???} and 4.4.2) and different stellar populations inside the bulge due to the merger (Section 5.3) with different $\Upsilon_{*}$ but overall bottom-heavy IMF (Section 5.1), which add to an overall rising $\Upsilon_\text{tot}(R')$ profile.}

>>>>

%--------------------------------------------
As mentioned above, JAM modelling allows to incorporate an invisible matter component in addition to the luminous matter in the form of an MGE. In Section \ref{sec:results_JAM_NFW} we will include a spherical Navarro-Frenk-White (NFW) DM halo \citep{1996ApJ...462..563N} in the dynamical model. The classical NFW profile has the form
\begin{equation}
\rho_\text{NFW}(r) \propto r^{-1} \left( r+r_s \right)^{-2} \label{eq:NFWprofile}
\end{equation}
and two free parameters, the scale length $r_s$ and a parameter describing the total mass of the halo. We will use $v_\text{200}\NEW{\equiv \sqrt{GM_{200}/r_{200}}}$, which is the circular velocity at the radius $r_\text{200}$ within which the mean density of the halo is 200 times the cosmological critical density $\rho_\text{crit} \equiv (2H^2)/(8\pi G)$, \OLD{i.e.,}\NEW{ and where $M_{200} \equiv M(<r_{200}) \stackrel{!}{=} \frac 43 \pi r_{200}^3 \cdot \rho_\text{crit}(z=0)$.}
\OLD{\begin{eqnarray*}
M_\text{200} &=& M(<r_{200})\\
\frac{M_{200}}{ \frac 43 \pi r_{200}^3} &=& 200 \rho_\text{crit}(z=0) \\
v_\text{200} &=& \sqrt{\frac{GM_{200}}{r_\text{200}}}
\end{eqnarray*}}
with $\rho_\text{crit}(z=0)=1.43 \cdot 10^{-7} M_\odot / \text{pc}^3$ in the WMAP5 cosmology by \citet{WMAP5cosm}. \OLD{How much the mass is concentrated in the center of the NFW halo is given by the concentration parameter
\begin{equation}
c_{200}\equiv r_{200} / r_s. \label{eq:NFW_c}
\end{equation}}
\NEW{The mass concentration of the NFW halo is quantified by $c_{200}\equiv r_{200} / r_s$.}
There is a close relation between the concentration and halo mass in simulations \citep{1996ApJ...462..563N}. \citet{Maccio08} found this relation for the WMAP5 cosmology to be
\begin{equation}
\langle \log c_{200} \rangle (M_{200}) = 0.830 - 0.098 \log \left(h \frac{M_{200}}{10^{12} M_\odot} \right) \label{eq:Maccio08}
\end{equation}
(their equation 10), with a Gaussian scatter of $\sigma_{\log c_{200}} = 0.105$ (their Table A2). \NEW{In Section \ref{sec:results_JAM_NFW} we will use this relation as prior information to guide the modelling towards realistic NFW halo shapes.}
%--------------------------------------------

@@

%--------------------------------------------
\subsubsection{Best fit JAM model} \label{sec:results_JAM_NFW_results}
%--------------------------------------------

@@

\NEW{A fit to all nine available $v_\text{rms}$ data points would force the DM halo to be less massive. But with decreasing DM contribution in the inner regions the model approaches the mass-follows-light model in the center, that we already ruled out in Section \ref{sec:results_JAM_SB_MfL}. Allowing for free $\Upsilon_{*}(R')$, $\beta_z(R')$ and/or non-NFW $\rho_\text{halo}(r)$ profiles would over-fit the data again. Without further priors on the stellar $\Upsilon_{*}(R')$ and/or more data points, we can make only the following statement: A model with constant $\Upsilon_*$, moderately tangential $\beta_z$ and a spherical and massive DM component with high concentration can explain the central kinematics of J1331 very well, we have however ruled out that in this case the DM halo will follow a NFW halo profile at both small and large radii.}
%--------------------------------------------

@@

%--------------------------------------------
\subsection{On J1331's possible merger history\OLD{ and modelling failures}} \label{sec:discussion_merger}
%--------------------------------------------
\NOTE{[This subsection was moved from a later position to the beginning of this section.]}

@@

\NEW{\emph{(i) Indications for a slightly less bottom-heavy Salpeter-like IMF in J1331's bulge:}}

@@

\NEW{This rise in $\Upsilon_\text{I,*}(R')$ could be either due to a cuspy DM halo (see Section \ref{sec:results_JAM_NFW}) or a very strong gradient in $\Upsilon_\text{I,*}$ (see Section \ref{sec:discussion_merger}). However, we rule out that a DM cusp is the sole reason because a cuspy and therefore overall massive DM halo does not match the kinematics in J1331's outer regions. We also rule out that a very strong $\Upsilon_\text{I,*}$ gradient is the only reason, because---as mentioned in Section \ref{sec:discussion_merger}---photometry did not reveal the clear existence of a blue population in the very center of J1331's bulge. Also, our modelling attempts allowing for velocity anisotropy (Sections \ref{sec:results_JAM_SB_MfL} and \ref{sec:results_JAM_NFW}) suggest that we do need some moderate $\beta_z<0$ to explain the central $v_\text{rms}$ dip. And lastly---as laid out in the previous paragraph---J1331's central velocity dispersion suggests a more bottom-heavy IMF. Overall it is therefore not very likely that the central regions of J1331 have such a low $\Upsilon_\text{I,*}^\text{chab} \sim 2.5$.}

@@

\NEW{The central $v_\text{rms}$ dip is most likely a combination of the three different aspects that we investigated separately: We expect a bit of tangential velocity anisotropy (Sections \ref{sec:results_JAM_SB_MfL} and \ref{sec:results_JAM_NFW}), a moderate contribution of a roundish DM halo (Sections \ref{sec:results_lensing_compare} and \ref{sec:results_JAM_NFW_results}) and different stellar populations inside the bulge due to the merger (Section \ref{sec:discussion_merger}) with different $\Upsilon_{*}$ but overall bottom-heavy IMF (Section \ref{sec:MLdiscussion}), which add to an overall rising $\Upsilon_\text{tot}(R')$ profile.}
